# Reproducible Research: Peer Assessment 1
========================================

Set your working directory.
Then create a subdirectory called data
# Create directory if not present
```{r, echo=TRUE}
if(!file.exists("data")){
  dir.create("data") }
```
## Download the file and extract to it to the data folder that was just created
Download the data file and save it in your working directory. 
The code below will unzip your zip file, place the contents into the data directory that was just created
and set your working directory to data.
The activity.zip file can be downloaded from [activity data set](https://github.com/rdpeng/RepData_PeerAssessment1/blob/master) 
```{r, echo=TRUE}
unzip(zipfile = "./activity.zip", exdir = "./data")
```

### Read the data in and transform the columns as needed. This includes converting the date column to dates, adding a column called intervalFactor in which the intervals are converted to factors. Create a column called DayNames using the weekdays function on the dates. This column now contains the names of the day of the week for each date. The sixth column is added to store whether the day is a weekday or a weekend. 
```{r, echo=TRUE}
setwd("./data")
activity <- read.csv("./activity.csv")
activity$Dates <- as.Date(as.character(activity$date))
activity$DayNames <- weekdays(activity$Dates)
activity[((activity$DayNames == "Saturday")|(activity$DayNames=="Sunday")),6] <- "weekend"
activity[(!(activity$DayNames == "Saturday")|(activity$DayNames=="Sunday")),6] <- "weekday"
##activity$intervalFactor <- factor(activity$interval)
```

### Create a histogram of the steps per day
The code below creates a histogram of the frequency of number of steps per day, and puts in a line for the median
and the mean number of steps per day as well
```{r, echo=TRUE}
#Calculate the steps per day
steps_per_day <- aggregate(steps ~ date, activity, sum)
hist(steps_per_day$steps, main="Frequency Of Steps Per Day", breaks = nrow(steps_per_day), ylim=c(0,8), xlab="Steps Per Day")
abline(v=mean(steps_per_day$steps), col="yellow", lwd=2.0, lty="dashed")
abline(v=median(steps_per_day$steps), col="red", lwd=1.0, lty="dotted")
legend("topright", lty=c(2,3), col=c("yellow", "red", "blue"), legend=c("mean steps per day","median steps per day"),cex=.7)
```


### What is mean total number of steps taken per day?
calculate the mean and median number of steps per day and report them
```{r, echo=TRUE}
meanspd <- round(mean(steps_per_day$steps),2)
medianspd <- round(median(steps_per_day$steps))
```

The mean number of steps per day is `r format(meanspd, scientific=F)`
The median number of steps per day is `r format(medianspd, scientific=F)`


## What is the average daily activity pattern?
The code below will create a new column from the interval data. The column class will be factor. This will 
be used for calculating the average for each five minute interval across all days
```{r, echo=TRUE}

w <- sprintf("%04i",activity$interval)
z <- strptime(w, "%H%M")
timeonly <- format(z, format="%H:%M:%S")
##activity$intervalFactor <- factor(timeonly)
activity$intervalFactor<-factor(activity$interval)
avg_per_interval <- aggregate(steps ~ intervalFactor, activity, mean, rm.na=T)

colnames(avg_per_interval) <- c("interval", "steps")
avg_per_interval$steps <- integer(avg_per_interval$steps)
plot(avg_per_interval, main = "Average Daily Activity Pattern", xlab = "5 Minute Intervals",
     ylab = "Avg steps taken per 5 minute interval")
lines(avg_per_interval$interval,avg_per_interval$steps, type="l")
most <- avg_per_interval$interval[which.max(avg_per_interval$steps)]
```
The interval during the time `r format(most, scientific=F)` had the most steps per interval

## Imputing missing values
Calculate the number of missing values in the data set
Use the average number of steps per interval to fill in the missing steps values
You need to install the Hmisc package for this method of imputation
```{r, echo=TRUE}
install.packages("Hmisc")
library(Hmisc)
t <- table(is.na(activity$steps))
numMissing <- t[2]
cleaning <- activity
orig <- activity$steps
stepsMeans <-iavg_per_interval$steps
for(i in )
nas_steps <- which(is.na(orig))
defaults <- steps_per_interval
    na_replacements <- unlist(lapply(na_indices, FUN=function(idx){
        interval = tbl[idx,]$interval
        defaults[defaults$interval == interval,]$steps
        }))
    imp_steps <- tbl$steps
    imp_steps[na_indices] <- na_replacements
if (is.na(orig)){
  steps <- subset(avg_per_interval, interval=cleaning$interval)$steps
}else{
    steps <-cleandata$steps
}
cleandata <- c(cleaning, steps)


```
There are `r format(numMissing, scientific = F)` rows with NA's in the data

## Are there differences in activity patterns between weekdays and weekends?
